---
title: "Coursera Reproducible Research Assignment"
output: html_document
---

## Step1
-Loading and preprocessing the data

## Step2 

- confirm that the imported data is indeed a dataframe
- add weekdays to the dataframe
- create weekday column
- create table to identify the counts per weekday
- create grouped values


```{r}
library(ggplot2)
library(zoo)
library(dplyr)
library(tidyverse)


readfile <- read.csv("activity.csv", header = TRUE, sep = ',', colClasses = c("numeric", "character", "integer"))
readfile$date <- as.Date(readfile$date)
readfile$weekdays <- weekdays(readfile$date)
head(readfile)
weekdays_count <- table(readfile$weekdays)
head(weekdays_count)
weekdays_count <- as.data.frame(weekdays_count)
```

## Step3
- What is mean total number of steps taken per day? +
- What is the average daily activity pattern?
- In the following histogramm we can see the total number of steps per day - accomplished via application of group_by and summarise functions.
- we also verify that the TotalSteps column has the class "integer"
- Below that we will find the median and the mean per day aswell as the mean and median for the entire column.
- Now, we will join the 3 new dataframes created previously in order to be able to access all variables more easily.
- We will also group the variables by interval now instead of dates and create an interactive plot using plotly
- The 835 Intervall has the highest value (206,17)

```{r}
totalstepsperday <- as.data.frame(readfile) %>% group_by(date) %>%  summarise(TotalSteps = sum(steps,na.rm = TRUE))

hist(totalstepsperday$TotalSteps)

head(totalstepsperday)

meantotalsteps <- mean(totalstepsperday$TotalSteps)

meantotalsteps

mediantotalsteps <- median(totalstepsperday$TotalSteps)

mediantotalsteps

totalstepsperdayMEAN <- readfile %>%  group_by(date) %>% summarise(TotalSteps = mean(steps,na.rm = TRUE))

  
totalstepsperdayMEDIAN <- readfile %>%   group_by(date) %>%  summarise(TotalSteps = median(steps,na.rm = TRUE))


combined_Data_frame <- right_join(totalstepsperdayMEAN,totalstepsperday, by=c("date"))

combined_Data_frame <- right_join(combined_Data_frame,totalstepsperdayMEDIAN, by=c("date"))

stepsperinterval <- readfile %>%  group_by(interval) %>%  summarise(MEAN_Steps = mean(steps,na.rm = TRUE))

max(stepsperinterval$MEAN_Steps,na.rm = TRUE)
  
stepsintervalplot <- ggplot(stepsperinterval, aes(x=interval, y    =MEAN_Steps)) + geom_line(color = "darkmagenta")

stepsintervalplot

```


## Step4

1) Calculate and report the total number of missing values


2) Devise a strategy for filling in all of the missing values in the dataset

I replaced all NAs with the average (for the entire dataset)


Probleme:


```{r}
NAcount <- sum(is.na(readfile))
NAcount

```

```{r}

plot<- ggplot(weekdays_count, aes(factor(Var1), Freq, fill = Var1)) +  geom_col(position = 'dodge')

readfileNAimputed <- as.data.frame(readfile)

readfileNAimputed$steps[is.na(readfileNAimputed$steps)] <- mean(readfileNAimputed$steps,na.rm=TRUE)


stepsperintervalbyday <- as.data.frame(readfile) %>% group_by(date) %>%  summarise(meansteps = mean(steps,na.rm = TRUE))
stepsperintervalbyday <- as.data.frame(stepsperintervalbyday)
stepsperintervalimputed <- readfileNAimputed %>%  group_by(date) %>%  summarise(meansteps = mean(steps,na.rm = TRUE))
 
hist(stepsperintervalbyday$meansteps,breaks=61)

hist(stepsperintervalimputed$meansteps,breaks=61)

data_full <- readfile

nas <- is.na(data_full$steps)

avg_interval <- tapply (data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)

data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]


sum(is.na(data_full$steps)) ## [1] 0

steps_full <- data_full %>% filter(!is.na(steps)) %>%   group_by(date) %>%  summarize(steps = sum(steps))

auxplot <- ggplot(steps_full, aes(x = steps)) + geom_histogram(fill = "firebrick", binwidth = 1000) + labs(title = "Histogram of Steps per day, including missing values", x ="Steps per day", y = "Frequency")


steps_full$steps <- as.numeric(steps_full$steps)

mean_steps_full <- mean(steps_full$steps,na.rm=TRUE)

medianstepsfull <- median(steps_full$steps, na.rm = TRUE)

mean_steps_full

medianstepsfull

```

## Step5
- Are there differences in activity patterns between weekdays and weekends?
- Yes, the person starts earlier on weekdays but does more during the weekend.
- A guess would be that the person works in an office and therefore does less steps when he works.


```{r}

data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Samstag" | weekdays(data_full$date) == "Sonntag", "weekend", "weekday"))

data_full$weektype <- as.factor(data_full$weektype)

interval_full <- data_full %>% group_by(interval, weektype) %>% summarise(steps = mean(steps))

ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) + geom_line() + facet_wrap(~weektype, ncol = 1, nrow=2)


```



